The non-random assembly of functional motifs in  \nplant-pollinator networks

### 2022-03-24 Manuscript/Main_text/Main_text_Functional_Ecology.Rmd
Project organization

---
title: "The non-random assembly of functional motifs in plant-pollinator networks worldwide"
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{caption,setspace}
- \captionsetup[figure]{labelformat=empty}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document: default
  html_document: default
  word_document: default
always_allow_html: true
bibliography: bibliography.bib
biblio-style: apalike
csl: oikos.csl #oikos style from zotero styles
---

@@

Understanding the global patterns of network structure is key to unravel the processes that govern community dynamics [@bascompte2007; @bastolla2009; @guimaraes2020]. However, common analytical approaches at the node or at the full network level are unable to capture the full complexity of ecological systems [@simmons2019; @bergamo2021]. Because network motifs can capture both direct and indirect interactions, by analysing over- and under-representation of motifs we are able to explore in great detail the structure of interactions of plant-pollinator networks. We have explored for the first time the overall structure and representation of network motifs in real plant-pollinator networks at a macro-ecological scale and shown that the different motifs up to five nodes are consistently over- or under-represented in comparison with simulated networks, therefore indicating the presence of ecological processes shaping their structure. Despite the presence of multiple ecological processes likely operating at the same time (e.g., neutral effects, morphological matching or phenological overlap), we found a tendency for over-represented structures with more densely connected motifs (i.e., complete, fan and asymmetric complete). As suggested in @simmons2020, a higher proportion of these densely connected motifs could reflect that neutral effects are the predominant ecological process governing the structure of these network motifs. Thus, as indicated for other network properties like nestedness at the macro-scale level [@krishna2008; @suweis2013], species relative abundances (i.e., neutral effects) may be also the predominant ecological process governing the meso-scale structure on plant-pollinator networks.

>>>>

---
title: "The non-random assembly of functional motifs in  \nplant-pollinator networks"
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{caption,setspace}
- \captionsetup[figure]{labelformat=empty}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  pdf_document: default
  html_document: default
  word_document: default
always_allow_html: true
bibliography: bibliography.bib
biblio-style: apalike
csl: oikos.csl #oikos style from zotero styles
---

@@

Understanding the global patterns of network structure is key to unravel the processes that govern community dynamics [@bascompte2007; @bastolla2009; @guimaraes2020]. However, common analytical approaches at the node or at the full network level are unable to capture the full complexity of ecological systems [@simmons2019; @bergamo2021]. Because network motifs can capture both direct and indirect interactions, by analysing over- and under-representation of motifs we are able to explore in great detail the structure of interactions of plant-pollinator networks. We have explored for the first time the overall structure and representation of network motifs in real plant-pollinator networks at a macro-ecological scale and shown that the different motifs up to five nodes are consistently over- or under-represented in comparison with simulated networks, therefore indicating the presence of ecological processes shaping their structure. Despite the presence of multiple ecological processes likely operating at the same time (e.g., neutral effects, morphological matching or phenological overlap), we found a tendency for over-represented structures with more densely connected motifs (i.e., complete, fan and asymmetric complete). As suggested in @simmons2020, a higher proportion of these densely connected motifs could reflect that neutral effects are the predominant ecological process governing the structure of network motifs. Thus, as indicated for other network properties like nestedness at the macro-scale level [@krishna2008; @suweis2013], species relative abundances (i.e., neutral effects) may be also the predominant ecological process governing the meso-scale structure on plant-pollinator networks.
